## 6A Lab replicates and Exercise

### 1.1 Regression Trees (Boston Housing Data)
```{r message=FALSE}
library(MASS) #this data is in MASS package
set.seed(14023799)
boston_data <- data(Boston)
sample_index <- sample(nrow(Boston),nrow(Boston)*0.90)
boston_train <- Boston[sample_index,]
boston_test <- Boston[-sample_index,]
```

```{r message=FALSE, warning=FALSE}
library(rpart)
library(rpart.plot)
```

```{r}
boston_rpart <- rpart(formula = medv ~ ., data = boston_train)
```


#### 1.1.1 Printing and ploting the tree
```{r}
boston_rpart
```
```{r}
prp(boston_rpart,digits = 4, extra = 1)
```


#### Exercise Q1 :Based on the tree model above, what is the predicted median housing price (in thousand) given following information:

**Ans : The predicted median housing price is 27.63**


+ In-sample prediction
```{r}
boston_train_pred_tree = predict(boston_rpart)
```

+ Out-of-sample prediction

```{r}
boston_test_pred_tree = predict(boston_rpart,boston_test)
```

#### Exercise Q2:  Calculate the mean squared error (MSE) of the in-sample for this tree model

```{r}
mean((predict(boston_rpart) - boston_train$medv)^2)
```
**Ans : The in-sample MSE for this tree model is 17.148**


```{r}
boston.reg = lm(medv~., data = boston_train)
boston_test_pred_reg = predict(boston.reg, boston_test)
mean((boston_test_pred_reg - boston_test$medv)^2)
```


```{r}
boston_test_pred_rpart = predict(boston_rpart, boston_test)
mean((boston_test_pred_rpart - boston_test$medv)^2)
```
#### Exercise Q3 : Compare this model’s out-of-sample performance with the linear regression model with all variables in it.

**Ans : The tree model's out-of-sample MSPE is 10.406 whereas that of linear regression with all models is 13.187. So the out-of-sample performance is better for tree model as it has lower MSPE value**


#### 1.2 Pruning

```{r}
boston_largetree <- rpart(formula = medv ~ ., data = boston_train, cp = 0.001)
```

```{r}
prp(boston_largetree)
```

```{r}
plotcp(boston_largetree)
```


```{r}
printcp(boston_largetree)
```

```{r}
sum((boston_train$medv - mean(boston_train$medv))^2)/nrow(boston_train)
```

```{r}
mean((predict(boston_largetree) - boston_train$medv)^2)
```

```{r}
prune(boston_largetree, cp = 0.008)
```

