## 6B Lab replicates and Exercise

#### 1 CART: classification and regression tree

##### 1.1 Classification Trees (Credit Card Default data)

```{r message=FALSE, warning=FALSE}
library(rpart)
library(rpart.plot)
set.seed(14023799)
```

```{r message=FALSE, warning=FALSE}
credit_data <- read.csv(file = "https://xiaoruizhu.github.io/Data-Mining-R/lecture/data/credit_default.csv", header=T)

# rename
library(dplyr)
```


```{r}
credit_data<- rename(credit_data, default=default.payment.next.month)
# convert categorical data to factor
credit_data$SEX<- as.factor(credit_data$SEX)
credit_data$EDUCATION<- as.factor(credit_data$EDUCATION)
credit_data$MARRIAGE<- as.factor(credit_data$MARRIAGE)

index <- sample(nrow(credit_data),nrow(credit_data)*0.80)
credit_train = credit_data[index,]
credit_test = credit_data[-index,]

credit_rpart0 <- rpart(formula = default ~ ., data = credit_train, method = "class")

credit_rpart <- rpart(formula = default ~ . , data = credit_train, method = "class", parms = list(loss=matrix(c(0,5,1,0), nrow = 2)))
```


```{r}
pred0 <- predict(credit_rpart0, type="class")
table(credit_train$default, pred0, dnn = c("True", "Pred"))
```


#### 1.1.1 Printing and ploting the tree

```{r}
credit_rpart
```

```{r}
prp(credit_rpart, extra = 1)
```


#### 1.1.2 Prediction using classification trees

+ In-sample prediction
```{r}
credit_train.pred.tree1<- predict(credit_rpart, credit_train, type="class")
table(credit_train$default, credit_train.pred.tree1, dnn=c("Truth","Predicted"))
```

### Exercise Q1: For the credit card default data, calculate the out-of-sample confusion matrix    

```{r}
#Predicted Class
credit_test.pred.tree1<- predict(credit_rpart, credit_test, type="class")
table(credit_test$default, credit_test.pred.tree1, dnn=c("Truth","Predicted"))
```



```{r}
cost <- function(r, phat){
  weight1 <- 5
  weight0 <- 1
  pcut <- weight0/(weight1+weight0) 
  c1 <- (r==1)&(phat<pcut) #logical vector - true if actual 1 but predict 0
  c0 <-(r==0)&(phat>pcut) #logical vector - true if actual 0 but predict 1
  return(mean(weight1*c1+weight0*c0))
}

cost(credit_train$default, predict(credit_rpart, credit_train, type="prob"))
```

#### 1.1.3 Calculate the actual cost using a self-defined cost function.

```{r}
cost(credit_test$default,  predict(credit_rpart, credit_test, type="prob"))
```

#### 1.1.4 Comparing this classification tree with logistic regression

```{r}
#Fit logistic regression model
credit_glm <- glm(default~., 
                  data = credit_train, 
                  family=binomial)
#Get binary prediction
credit_test_pred_glm <- predict(credit_glm, credit_test, type="response")
#Calculate cost using test set
cost(credit_test$default, credit_test_pred_glm)
```

```{r}
#Confusion matrix
table(credit_test$default, as.numeric(credit_test_pred_glm>1/6), dnn=c("Truth","Predicted"))
```


```{r}
#Get binary prediction
credit_train_pred_glm <- predict(credit_glm, credit_train, type="response")
#Calculate cost using test set
cost(credit_train$default, credit_train_pred_glm)
```
### Exercise Q2: If the cost ratio of false negative to false positive is 5 to 1, calculate the out-of-sample asymmetric cost.

```{r}
credit_test.pred.tree1<- predict(credit_rpart, credit_test, type="prob")
cost(credit_test$default, credit_test.pred.tree1)
```

**The out-of-sample asymmetric cost is 0.673**

#### 1.1.5 ROC Curve and Cut-off Probability for Classification trees
```{r}
credit_rpart <- rpart(formula = default ~ .,
                      data = credit_train, 
                      method = "class",
                      parms = list(loss=matrix(c(0,5,1,0), nrow = 2)))
#Probability of getting 1
credit_test_prob_rpart = predict(credit_rpart, credit_test, type="prob")
```

```{r}
library(ROCR)
```

```{r}
pred = prediction(credit_test_prob_rpart[,2], credit_test$default)
perf = performance(pred, "tpr", "fpr")
plot(perf, colorize=TRUE)
```

```{r}
slot(performance(pred, "auc"), "y.values")[[1]]
```
```{r}
credit_test_pred_rpart = as.numeric(credit_test_prob_rpart[,2] > 1/(5+1))
table(credit_test$default, credit_test_pred_rpart, dnn=c("Truth","Predicted"))
```

#### Exercise: Draw the ROC curve for training sample.

```{r}
credit_train_prob_rpart = predict(credit_rpart, credit_train, type="prob")
pred = prediction(credit_train_prob_rpart[,2], credit_train$default)
perf = performance(pred, "tpr", "fpr")
plot(perf, colorize=TRUE)
```



