### Data Mining II
### Yamini Tawde

## 7B Random forests
## Advanced Tree Models – Random Forests

```{r}
library(MASS)
data(Boston)
set.seed(14023799)
index <- sample(nrow(Boston),nrow(Boston)*0.9)
boston_train <- Boston[index,]
boston_test <- Boston[-index,]
```

**1 Random Forests**

**1.1 Random Forest for Regression**

```{r}
library(randomForest)
boston_rf<- randomForest(medv~., data = boston_train, importance=TRUE)
boston_rf
```

```{r}
mod_rf <- randomForest(medv~., data=boston_train, 
                       importance=TRUE, ntree=500)
```

```{r}
boston_rf$importance
```

```{r}
mod_rf$importance
```

```{r}
plot(boston_rf$mse, type='l', col=2, lwd=2, xlab = "ntree", ylab = "OOB Error")
```

```{r}
plot(mod_rf$mse, type='l', col=2, lwd=2, xlab = "ntree", ylab = "OOB Error")
```

**Prediction on the testing sample**
```{r}
boston_rf_pred<- predict(boston_rf, boston_test)
mean((boston_test$medv-boston_rf_pred)^2)
```

## Exercise 7B 

Apply random forests on the Boston housing data
Draw this type of plot below, the MSE of random forests with different (mtry) number of variables randomly sampled as candidates at each split.

```{r}
oob_err <- rep(0, 13)
test.err <- rep(0, 13)
for(i in 1:13){
  fit<- randomForest(medv~., data = boston_train, mtry=i)
  oob_err[i]<- fit$mse[500]
  test.err[i]<- mean((boston_test$medv-predict(fit, boston_test))^2)
  cat(i, " ")
}
```
```{r}
matplot(cbind(test.err, oob_err), pch=15, col = c("red", "blue"), type = "b", ylab = "MSE", xlab = "mtry")
legend("topright", legend = c("test Error", "OOB Error"), pch = 15, col = c("red", "blue"))
```

**1.2 Random Forest for Classification**


```{r}
# load credit card data
credit_data <- read.csv(file = "https://xiaoruizhu.github.io/Data-Mining-R/lecture/data/credit_default.csv", header=T)
# convert categorical variables
credit_data$SEX<- as.factor(credit_data$SEX)
credit_data$EDUCATION<- as.factor(credit_data$EDUCATION)
credit_data$MARRIAGE<- as.factor(credit_data$MARRIAGE)
# random splitting
set.seed(14023799)
index <- sample(nrow(credit_data),nrow(credit_data)*0.9)
credit_train = credit_data[index,]
credit_test = credit_data[-index,]

credit_rf <- randomForest(as.factor(default.payment.next.month)~., 
                          data = credit_train,
                          importance=TRUE, ntree=500)
credit_rf
```

```{r}
plot(credit_rf, lwd=rep(2, 3))
legend("right", legend = c("OOB Error", "FPR", "FNR"), lwd=rep(2, 3), lty = c(1,2,3), col = c("black", "red", "green"))
```
**ROC curve and AUC can be obtained based on the probability prediction**
```{r}
credit_rf_pred <- predict(credit_rf, type = "prob")[,2]
library(ROCR)
pred <- prediction(credit_rf_pred, credit_train$default.payment.next.month)
perf <- performance(pred, "tpr", "fpr")
plot(perf, colorize=TRUE)
```
```{r}
#Get the AUC
unlist(slot(performance(pred, "auc"), "y.values"))
```
**Create the confusion matrix based on the cutoff probability from asymmetric cost (pcut=1/6)**
```{r}
## out-of-sample
pcut <- 1/6
credit_rf_pred_test <- predict(credit_rf, newdata=credit_test, type = "prob")[,2]
credit_rf_class_test <- (credit_rf_pred_test>pcut)*1
table(credit_test$default.payment.next.month, credit_rf_class_test, dnn = c("True", "Pred"))
```

