
## Generalized Additive Model

### 1 Generalized Additive Model

**1.1 GAM on Boston Housing dataset**

```{r}
library(MASS)
set.seed(14023799)
sample_index <- sample(nrow(Boston),nrow(Boston)*0.70)
Boston_train <- Boston[sample_index,]
Boston_test <- Boston[-sample_index,]
str(Boston_train)
```

```{r}
library(mgcv)

#create gam model
Boston_gam <- gam(medv ~ s(crim)+s(zn)+s(indus)+chas+s(nox)
                 +s(rm)+s(age)+s(dis)+rad+s(tax)+s(ptratio)
                 +s(black)+s(lstat),data=Boston_train)

summary(Boston_gam)
```

```{r}
plot(Boston_gam, pages=1)
```

**Model AIC/BIC and mean residual deviance**
```{r}
AIC(Boston_gam)
BIC(Boston_gam)
Boston_gam$deviance
```

**In-sample fit performance**

```{r}
#in-sample mse using df 
Boston_gam.mse.train <- Boston_gam$dev/Boston_gam$df.residual 
#Average Sum of Squared Error
Boston_gam.mse.train <- Boston_gam$dev/nrow(Boston_train) 

#using the predict() function
pi <- predict(Boston_gam,Boston_train)
mean((pi - Boston_train$medv)^2)
```

```{r}
pi.out <- predict(Boston_gam,Boston_test)
mean((pi.out - Boston_test$medv)^2)
```

### 1.2 GAM on Bankruptcy dataset
```{r}
Bank_data <- read.csv(file = "https://xiaoruizhu.github.io/Data-Mining-R/lecture/data/bankruptcy.csv", header=T)
# summary(Bank_data)

sample_index <- sample(nrow(Bank_data),nrow(Bank_data)*0.70)
Bank_train <- Bank_data[sample_index,]
Bank_test <- Bank_data[-sample_index,]
```


```{r}
Bank_gam <- gam(DLRSN ~ s(R1)+s(R2)+s(R3)+s(R4)+
                  s(R5)+s(R6)+s(R7)+s(R8)+s(R9)+s(R10), data=Bank_train)

summary(Bank_gam)
```

```{r}
plot(Bank_gam, pages=1)
```

**In-sample fit performance**
```{r}
pcut_gam <- 1/36
prob_gam_in <-predict(Bank_gam, Bank_train, type="response")
pred_gam_in <- (prob_gam_in>=pcut_gam)*1
table(Bank_train$DLRSN, pred_gam_in, dnn=c("Observed","Predicted"))
```


```{r}
bankcost <- function(r, pi){
  weight1 = 35
  weight0 = 1
  pcut <- weight0/(weight0+weight1)
  c1 = (r==1)&(pi<pcut) #logical vector - true if actual 1 but predict 0
  c0 = (r==0)&(pi>pcut) #logical vector - true if actual 0 but predict 1
  return(mean(weight1*c1+weight0*c0))
}
bankcost(Bank_train$DLRSN, pred_gam_in)
```

**ROC Curve:**

```{r}
library(ROCR)
pred <- prediction(c(prob_gam_in), Bank_train$DLRSN)
perf <- performance(pred, "tpr", "fpr")
plot(perf, colorize=TRUE)
```


```{r}
#Get the AUC
unlist(slot(performance(pred, "auc"), "y.values"))
```

**Model AIC/BIC and mean residual deviance**

```{r}
AIC(Bank_gam)
```

```{r}
BIC(Bank_gam)
```

```{r}
#in-sample mean residual deviance using df 
Bank_gam$dev/Bank_gam$df.residual 
```



**Out-of-sample fit performance**

```{r}
prob_gam_out <- predict(Bank_gam, Bank_test, type="response")
pred_gam_out <- (prob_gam_out>=pcut_gam)*1
table(Bank_test$DLRSN, pred_gam_out, dnn=c("Observed","Predicted"))
```


**The asymmetric cost is:**
```{r}
bankcost(Bank_test$DLRSN, pred_gam_out)
```

**ROC Curve:**
```{r}
pred <- prediction(c(prob_gam_out), Bank_test$DLRSN)
perf <- performance(pred, "tpr", "fpr")
plot(perf, colorize=TRUE)
```

```{r}
#Get the AUC
unlist(slot(performance(pred, "auc"), "y.values"))
```

### 1.3 GAM on Credit Card Default Data

```{r message=FALSE, warning=FALSE}
credit_data <- read.csv(file = "https://xiaoruizhu.github.io/Data-Mining-R/lecture/data/credit_default.csv", header=T)

# rename
library(dplyr)
credit_data<- rename(credit_data, default=default.payment.next.month)
# convert categorical data to factor
credit_data$SEX <- as.factor(credit_data$SEX)
credit_data$EDUCATION <- as.factor(credit_data$EDUCATION)
credit_data$MARRIAGE <- as.factor(credit_data$MARRIAGE)
```

```{r}
index <- sample(nrow(credit_data),nrow(credit_data)*0.90)
credit_train = credit_data[index,]
credit_test = credit_data[-index,]
```

```{r}
## Create a formula for a model with a large number of variables:
gam_formula <- as.formula("default~s(LIMIT_BAL)+s(AGE)+s(PAY_0)+s(BILL_AMT1)+s(PAY_AMT1)+SEX+EDUCATION+MARRIAGE")

credit_gam <- gam(formula = gam_formula, family=binomial, data=credit_train);
summary(credit_gam)
```

```{r}
plot(credit_gam, shade=TRUE, seWithMean=TRUE, scale=0, pages = 1)
```

```{r}
# vis.gam(credit_gam)
vis.gam(credit_gam, view=c("LIMIT_BAL","AGE"), theta= 140) # different view 
```


## Exercise 8B


**For the credit scoring data with GAM example, is the LIMIT_BAL variable nonlinearly associated with the default status?
What about the AGE?**

+ Yes the LIMIT_BAL variable is nonlinearly associated with default status  as the edf value is 5.900
+ The AGE is linearly associated as the edf value is 1 and also same conclusion can be drawn from above  vis.gam graph


**1.3.1 In-sample fit performance**

```{r}
pcut_gam <- 1/6
prob_gam_in <-predict(credit_gam,credit_train,type="response")
pred_gam_in <- (prob_gam_in>=pcut_gam)*1
table(credit_train$default, pred_gam_in,dnn=c("Observed","Predicted"))
```

```{r}
creditcost <- function(r, pi){
  weight1 = 5
  weight0 = 1
  pcut <- weight0/(weight0+weight1)
  c1 = (r==1)&(pi<pcut) #logical vector - true if actual 1 but predict 0
  c0 = (r==0)&(pi>pcut) #logical vector - true if actual 0 but predict 1
  return(mean(weight1*c1+weight0*c0))
}
creditcost(credit_train$default, pred_gam_in)
```


```{r}
AIC(credit_gam)
```

```{r}
BIC(credit_gam)
```

**ROC Curve**

```{r}
library(ROCR)
pred <- prediction(predictions = c(prob_gam_in), labels = credit_train$default)
perf <- performance(pred, "tpr", "fpr")
plot(perf, colorize=TRUE)
```

```{r}
#Get the AUC
unlist(slot(performance(pred, "auc"), "y.values"))
```

**1.3.2 Out-of-sample fit performance**

```{r}
pcut <- 1/6
prob_gam_out <- predict(credit_gam, credit_test,type="response")
pred_gam_out <- (prob_gam_out>=pcut)*1
table(credit_test$default, pred_gam_out,dnn=c("Observed","Predicted"))
```

```{r}
creditcost(credit_test$default, pred_gam_out)
```

```{r}
pred <- prediction(predictions = c(prob_gam_out), labels = credit_test$default)
perf <- performance(pred, "tpr", "fpr")
plot(perf, colorize=TRUE)
```

```{r}
unlist(slot(performance(pred, "auc"), "y.values"))
```

### 1.4 GAM using the “Motorcycle” dataset

```{r}
library(MASS)
data('mcycle')
str(mcycle)
summary(mcycle)
```

```{r}
# Rename the variables for ease of usage
Y <- mcycle$accel
X <- mcycle$times

#Scatterplot
plot(Y~X, xlab="time",ylab="Acceleration", main="Scatterplot of Acceleration against Time")
```

```{r}
library(mgcv)
s_gam <- gam(Y ~ s(X),data=mcycle)
summary(s_gam)
```


```{r}
#plot the model
plot(s_gam, residuals = TRUE, pch = 1)
```

